How Slow do my Transient Typechecks Go?

### 2019-08-17 paper.tex
J done 4

\begin{figure*}
	\input{graphs/full-scatter}
	\caption{Graphs of (at most) 102 configurations in the typing latices for each benchmark. Time is measured as the mean of the 351\textsuperscript{st} to the 1,000\textsuperscript{th} benchmark iteration under a single invocation of Moth (lower is better).}
	\label{f:full}
\end{figure*}
The results for the performance of our sample of typing lattice configurations are presented in figure~\ref{f:full}. Following \cite{vitousek-transient-arXive-2019}, the points on each graph show the average execution if each individual configuration. The x-axis represents the proportion of type annotations for each configuration, with the left and right most points showing the times for the fully untyped and typed configurations (respective). The execution time (in milliseconds) is shown on the left y-axes, and time relative to the fully untyped configuration is shown on the right y-axes.

Most of the graphs are mostly horizontal lines: indicating that the performance effect of including type annotations is negligible. However the plots for CD and Richards show a roughly linear increase: i.e. adding more type annotations to the program increases performance linearly. The plots for Go, Permute, DeltaBlue, and Storage even show \emph{decreases}: i.e. adding more type annotations can actually \emph{increase} performance. Of particular note is that some of the graphs appear to show bimodal performance profiles (Permute, Storage, Snake, Towers, and Richards), or in the case of List, trimodal.

>>>>

\begin{figure*}
	\input{graphs/full-scatter}
	\caption{Graphs of (at most) 102 configurations in the typing latices for each benchmark. Time is measured as the mean of the 351\textsuperscript{st} to the 1,000\textsuperscript{th} benchmark iteration under a single invocation of Moth (lower is better).}
	\label{f:full}
\end{figure*}
The results for the performance of our sample of typing lattice configurations are presented in figure~\ref{f:full}. Following \cite{vitousek-transient-arXive-2019}, the points on each graph show the average execution of each individual configuration. The x-axis represents the proportion of type annotations for each configuration, with the left and right most points showing the times for the fully untyped and typed configurations respectively. The execution time in milliseconds is shown on the left y-axes, and time relative to the fully untyped configuration is shown on the right y-axes.

Most of the graphs are essentially horizontal lines, indicating that the overhead of including type annotations is negligible. The plots for CD and Richards show a roughly linear increase, however, i.e.\ for these two benchmarks, adding type annotations increases performance linearly. On the other hand, the plots for Go, Permute, DeltaBlue, and Storage show \emph{decreases}: i.e. adding more type annotations \emph{increases} performance of these benchmarks.

Of particular note is that some of the graphs (Permute, Storage, Snake, Towers, and Richards) appear to show bimodal performance profiles, that is, two separate roughly horizontal lines. Presumably Moth can remove all the overhead from some configurations, but in others there must be one or more type checks that cannot be optimised away. List appears to have three performance modes: most of the configurations up to about 60\% annotations perform at roughly the same speed as untyped code; most configurations above 70\% perform 1.8 times more slowly than untyped code; and then across all annotation percentages there are configurations that perform roughly 1.5 times slower than the untyped baseline.
